---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->
<!-- The code to render this README is stored in .github/workflows/render-readme.yaml -->
<!-- Variables marked with double curly braces will be transformed beforehand: -->
<!-- `packagename` is extracted from the DESCRIPTION file -->
<!-- `gh_repo` is extracted via a special environment variable in GitHub Actions -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {{ packagename }}

{{ packagename }} provides functions to estimate one and two parameter functions
via maximum likelihood

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![R-CMD-check](https://github.com/{{ gh_repo }}/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/{{ gh_repo }}/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/{{ gh_repo }}/branch/main/graph/badge.svg)](https://app.codecov.io/gh/{{ gh_repo }}?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


## Installation

You can install the development version of {{ packagename }} from
[GitHub](https://github.com/) with:

```{r}
# check whether {pak} is installed
if(!require("pak")) install.packages("pak")
pak::pak("epiverse-trace/quickfit")
```

## Quick start

These examples illustrate some of the current functionalities:

```{r}
library(quickfit)
```

Generate some simulated data, define a likelihood, then estimate MLE, 
or MLE and 95% confidence interval based on profile likelihood:

```{r}
sim_data <- rnorm(50, 4, 2)

# Define likelihood function
log_l <- function(x,a,b) dnorm(x, a, b, log = TRUE)

# Estimate MLE
estimate_mle(log_l, sim_data, n_param = 2, a_initial = 3, b_initial = 1)

# Estimate 95% CI based on profile likelihood
calculate_profile(
  log_l, 
  data_in = sim_data, 
  n_param = 2, 
  a_initial = 3, 
  b_initial = 1, 
  precision = 0.01
)
```

Additionally, multiple distribution models can be compared (for censored and 
non-censored data).

```{r}
multi_fitdist(
  data = rlnorm(n = 100, meanlog = 1, sdlog = 1), 
  models = c("lnorm", "gamma", "weibull")
)
```

## Help 

To report a bug please open an [issue](https://github.com/epiverse-trace/quickfit/issues/new/choose)

### Contributions

Contributions are welcome via [pull requests](https://github.com/{{ gh_repo }}/pulls).

### Code of Conduct

Please note that the {{ packagename }} project is released with a 
[Contributor Code of Conduct](https://github.com/epiverse-trace/.github/blob/main/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
